# Caribou Analysis

```{r}
#| message: false

# library
library(tidyverse)
```

```{r}
# loading the data
load(here::here("data/caribou/caribou.Rdata"))

# data overview
glimpse(individuals)
glimpse(locations)

visdat::vis_dat(individuals)
visdat::vis_dat(locations, warn_large_data = FALSE)
```

The variable that does not contain NA in the individuals data are sex and study_site.

```{r}
# Joining the data
caribou <- individuals |> 
    select(animal_id, sex) |> 
    distinct(animal_id, .keep_all = TRUE) |> 
    right_join(locations, by = "animal_id") |> 
    mutate(
        sex = factor(sex, levels = c("f", "m"), labels = c("female", "male"))
    )
```


## Question

- Temporal migration patterns: Analyze how caribou movement changes throughout seasons and years.
- Herd comparison: Compare movement patterns between different herds to identify similarities and differences in behavior.
- Individual variation: Analyze differences in movement patterns between individual caribou within the same herd.
- Habitat preference: Identify areas where caribou spend more time versus areas they quickly pass through.
- Human impact assessment: Explore how proximity to human infrastructure (roads, settlements) affects caribou movement.
