# Network Analysis Experimentation

## 27th Feb 2025

The toy dataset structure:

**nodes**

- id
- lat
- lon
- type: either age_care or hospital
- size

**edges**

- from
- to
- timestamp
- reason: chest pain, breathing issue, injury, stroke, seizures, and high fever (from ChatGPT)


```{r}
#| results: false
# Loading the dataset
load(here::here("data/data/sim_network.RData"))
```

```{r}
head(edges)
head(nodes)
```

### Trying geomnet package

The data structure for the `geomnet` package combines two sources of information, node description and the relationship between nodes (vertices), into one dataset using the fortify methods. The supported classes are "network" from `sna` and `network` packages, "igraph" from the `igraph` package, "adjmat", and "edgedf" (classes from `geomnet`).

The `as.edgedf()` only works with data.frame object.

```{r}
#| message: false
library(dplyr)
edges <- edges |>
    mutate(from = factor(from, levels = seq(1:20)),
            to = factor(to, levels = seq(1:20)))
nodes <- nodes |>
    mutate(id = factor(id, levels = seq(1:20)))

edges_data <- as.data.frame(edges)
nodes_data <- as.data.frame(nodes)
```

Combining the edges and nodes

```{r}
#| message: false
#| output: false
library(geomnet)

network_data <- fortify(as.edgedf(edges_data), nodes_data)
head(network_data)
```

```{r}
#| eval: false
# Not working
set.seed(1)
library(ggplot2)

ggplot(data = network_data, 
        aes(from_id = from_id, to_id = to_id)) +
    geom_net()
```

There is a problem in the geom_net function where it said, "The first two columns of `x` must be of the same type." It also happens with the provided code in the `geomnet` package. Still not sure why.

https://rdrr.io/github/statnet/network/src/R/dataframe.R


## 6th Mar 2025

### tidygraph

The `tidygraph` provides a tidy API for graph manipulation. The network data can be seen as two tidy tables, one for node data and one for edge data. This package provides a way to switch between the two tables and provides dplyr verbs for manipulating the data. A lot of graph algorithms in `igraph` is also accessible through this package.

```{r}
library(tidygraph)

tidygraph_structure <- tbl_graph(nodes, edges)
```

#### Example

Change the size to a thousand. It is somewhat similar to `cubble` where I must activate the table I want to manipulate.

```{r}
tidygraph_structure |>
    activate(nodes) |>
    mutate(size = size/1000) |>
    activate(edges) |>
    mutate(reason = toupper(reason))
```


### sfnetworks

The `sfnetworks` allow us to extend the graph into the domain of geospatial data analysis, where each observation has a location in geographical space. The `sf` package is the driving force for this package. It is one of the common packages used when working with geospatial data.

```{r}
library(sf)
library(sfnetworks)

nodes_sf <- st_as_sf(nodes, coords = c("lat", "lon"))

sfnetwork_obj <- sfnetwork(nodes_sf, edges)
```

#### My first network plot

```{r}
library(ggplot2)

autoplot(sfnetwork_obj)
```